\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{float} % Table placement
\usepackage{bbm}

\title{Combinatorial Decision Making And Optimization}
\author{
  Cirone Cono, \texttt{cono.cirone@studio.unibo.it}
  \and
  Dardini Jacopo, \texttt{jacopo.dardini@studio.unibo.it}
  \and
  Formichella Gio, \texttt{gio.formichella@studio.unibo.it}
  \and
  Petrozziello Giulio, \texttt{giulio.petrozziello@studio.unibo.it} 
}
\date{}

\begin{document}

\maketitle


\section{Introduction}
In this project, we address the sports tournament scheduling (STS) problem by applying a unified modeling approach, inspired by \cite{10.1007/10704567_6}, across four computational paradigms: Constraint Programming (CP), Boolean Satisfiability (SAT), Satisfiability Modulo Theory (SMT) and Mixed Integer Programming (MIP).
We begin by considering the structure of a round-robin tournament: for $N$ teams, the tournament spans $N-1$ weeks, with each week consisting of $\frac{N}{2}$ games, one for each period. This structure naturally corresponds to a 1-factorization of the complete graph $K_N$ on $N$ vertices \cite{Dinitz2006RoundRobin}. Given this structure, all four approaches begin by precomputing the round-robin tournament matrix using a standard 1-factorization technique, yielding a $\frac{N}{2} \times (N-1)$ matrix where each column corresponds to a week and contains the set of weekly matchups. Our experimental results show that the time required to precompute the underlying round-robin schedule is negligible (less than a second) for all considered instance sizes. Precomputing the matrix provides several advantages:

\begin{itemize}
  \item It removes symmetries across weeks and matchup pairings, reducing the search space.
  \item It guarantees that each team plays exactly once per week.
  \item It ensures that each team plays against each other only once.
\end{itemize}
A balanced tournament design is a permutation of the order of games within each week that balances the number of home and away games for each team, and limits the number of times each team plays in the same period, as described in \cite{Lamken2006BalancedTournamentDesigns}. While the round-robin matrix defines the matchups, our models aim to compute an optimally balanced schedule by assigning these games to periods and slots in a way that satisfies the balancing constraints.

\subsection{Notation}
\begin{itemize}
  \item $N$: number of teams
  \item $T=\{t\ |\ t \in [1, N]\}$: team identifiers
  \item $P=\{p\ |\ p \in [0, ..., \frac{N}{2} - 1]\}$: period identifiers.
  \item $M=\{m\ |\ m \in [0, ..., \frac{N}{2} - 1]\}$: weekly matchup identifiers.
  \item $W=\{w\ |\ w \in [0, ..., N-2]\}$: week identifiers.
  \item $S=\{s\ |\ s \in [0,1]\}$: slot identifiers, $s=0$ corresponds to playing at home and $s=1$ to playing away.
  \item $rb$ is the precomputed round-robin matrix; $rb_{m, w, s}=t$ if team $t \in T$ plays in week $w \in W$ in match $m \in M$ and in slot $s \in S$.
  
\end{itemize}

\input{sections/CP}

\input{sections/SAT}

\input{sections/SMT}

\input{sections/MIP}

\input{sections/conclusion}

\bibliographystyle{unsrt}
\bibliography{ref.bib}

\end{document}
